---
title: "Main"
format: pdf
editor: visual
---

## Todos:

# 1) Briefly **describe** the data set (e.g., how many customers, what information are available, how do important variables look like).

\

\

```{r}
library(ggplot2)
library(ggcorrplot)
library(ggpubr)
library(rpart)
library(partykit)
library(dplyr)
library(gridExtra)
```

```{r}
set.seed(1234)
```

## Running Code

```{r}
load("Group1_streaming_ds.rda")
head(streaming_ds)
str(streaming_ds)
summary(streaming_ds)

streaming_ds["few hrs"] <- as.factor(streaming_ds$`few hrs`) # change column from boolean to factor 
```

# 2. **Identify the relevant variables** that would help you to segment/cluster/classify the users/customers and justify your choice. 

## Correlation Matrix

```{r}
streaming_only_num <- streaming_ds[sapply(streaming_ds, is.numeric)] # only get columns that are numberic

ggcorrplot(cor(streaming_only_num),
           lab = TRUE,               # Show correlation values
           lab_size = 3,             # Fix text size
           digits = 2,               # Number of decimal places
           method = "square")       # Square cells
```

No strong correlation for any of the numeric data points. However we have a few combinations which might be intersting to investigate further. Mainly: Income\~Age; Months\~Age (kind of obvious); Months\~Stopped

### Quick Viz for Columns against each other

```{r}
pairs(streaming_ds)
```

## Scatter plot for Income and age

```{r}
Age_income <- ggplot(streaming_ds, aes(x= age, y= income)) + 
  geom_point() + 
  scale_x_continuous(breaks = c(0,5,10,15,20,25,30,35,40,45,50,55,60,65,70,75,80,85))


Months_age <- ggplot(streaming_ds, aes(x= age, y= months)) + 
  geom_point() + 
  scale_x_continuous(breaks = c(0,5,10,15,20,25,30,35,40,45,50,55,60,65,70,75,80,85))

Months_stopped <- ggplot(streaming_ds, aes(x= months, y= stopped)) + 
  geom_point() + 
  scale_x_continuous(breaks= c(6, 12, 18, 24, 30, 36, 42))

Age_stopped <- ggplot(streaming_ds, aes(x=age, y = stopped)) + 
  geom_point() + 
  scale_x_continuous(breaks = c(0,5,10,15,20,25,30,35,40,45,50,55,60,65,70,75,80,85))


ggarrange(Age_income, Months_age, Months_stopped, Age_stopped, labels = c("Age ~ Income", "Months ~ Age", "Stopped ~ Months", "Age ~ stopped"), font.label = list(size=10), col= 1, nrow = 4)
```

Note: From the Months \~ Age plot it seem like Age and Months are not both describing the Age of the User, since this would mean a linear relationship

-\> Months is maybe the time of the subscription

### Income and Stopped

hallo

```{r}
ggplot(streaming_ds, aes(income, stopped)) +
  geom_point()
```

## Using a decision tree

### Pruning

```{r}
full_tree <- rpart(stopped ~. , data= streaming_ds, control = list(cp=0.001))

printcp(full_tree)
rpart::plotcp(full_tree)
```

### Pruned tree

From the decision tree it seems that the most important variables are the few_hrs, months, and age.

```{r}
pruned_tree <- prune(full_tree, cp = full_tree$cptable[5, "CP"])
plot(as.party(pruned_tree))
```

# 3) **Test** several models and **compare** them. **Consider the best approach** for your problem - clustering, classification, anything else?

-\> Predict if a customer is a Premium member or not Additionally predict number of stopped streams

Logistic Regression

# 4) **Justify your selection** of customers / respondents - what makes the group(s) special?

-\> Clustering to find likely customers that are not yet premium

```{r}
clustering_data <- streaming_ds %>%
  select(stopped, age, income, premium, months, `few hrs`) %>% # select relvant columns for sub df
  mutate(few_hrs = as.numeric(`few hrs`)) %>% # transform few hrs to numeric 
  select(-`few hrs`) %>% # remove old few hrs column 
  scale() %>% # scale for clustering
  as.data.frame(clustering_data) %>%
  `colnames<-`(c("stopped", "age", "income", "premium", "months", "few_hrs"))  # Assign column names


str(clustering_data)
```

```{r}
kmeans_results <- kmeans(clustering_data, centers = 2, nstart = 10)

clustering_df$Cluster <- as.factor(kmeans_results$cluster)

```

```{r}
p2 <- ggplot(clustering_df, aes(x = income, y=months, color = Cluster)) +
  geom_point(alpha = 0.6) +
  theme_minimal() +
  labs(title = "Clusters: Months vs Income")

p3 <- ggplot(clustering_df, aes(x = age, y = income, color = Cluster)) +
  geom_point(alpha = 0.6) +
  theme_minimal() +
  labs(title = "Clusters: Age vs Income")

```

```{r}
grid.arrange(p2, p3, ncol= 2)
```

```{r}
ggplot(clustering_data, aes(x = months, y = income, fill = factor(premium))) +
  geom_point(shape = 23, alpha = 0.6, size = 3) +
  scale_fill_manual(values = c("0" = "Blue", "1" = "Red"),
                    labels = c("Non-Premium", "Premium"),
                    name = "Subscription Status") +
  labs(
    title = "Income Distribution by Subscription Duration",
    x = "Months of Subscription",
    y = "Income"
  ) +
  theme_minimal()
```

# 5) Do the data reveal **how do the chosen customers differ** from the others?
